Here's a comprehensive R-based RNA-seq analysis pipeline using the `DESeq2` and
`edgeR` packages, suitable for small-to-mid-sized project work. This pipeline 
covers data loading, QC, alignment, count analysis, differential expression, 
visualization, and output visualization.

---

### 1. Installation and Setup

```{r ch1}
# Install required packages (install options may depend on OS and version)
install.packages(c("DESeq2", "tidyr", "dplyr", "ggplot2", "biocManager"))

# Load packages
library(DESeq2)
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggplot2)
```

---

### 2. Data Loading

```{r ch2}
# Example: Load counts and sample metadata
counts <- read.csv("counts.csv", header = TRUE)
metadata <- read.csv("metadata.csv", header = TRUE)
```

---

### 3. Basic QC and Quality Filtering

```{r ch3}
# Basic QC checks
sum(counts >= .001)  # Minimum expression
unique(counts)       # Sample names
```

---

### 4. Differential Expression with DESeq2

```{r ch4}
# Create sample information structure
design <- ~condition
dds <- DESeqDataSetFromMatrix(
  counts = counts,
  colData = metadata,
  design = design
)

dds <- DESeq(dds)

# Summary of results
results <- results(dds)  # or rownames(results) <- rownames(dds)

# View top changes
head(top50(results$table = TRUE))
```

---

### 5. Principal Component Analysis (PCA)

```{r ch5}
# PCA with PCA function
pca <- plotPCA(dds, intgroup = "condition")
```

---

### 6. Visualization of Differences

```{r ch6}
# Differential expression barplot
# Use ggplot2 or base R for bar plots
barplot(log2(abs(results$counts)), pch = 19, cex = 0.5)
abline(h = 0, lty = 2)
```

---

### 7. Additional Tools

```{r ch7}
# Example of using DESeq2 for filtering low-expression genes
dds <- DESeq(dds)
res <- results(dds)

# Filter genes by log2 fold change
res_filtered <- res[abs(res$log2FoldChange) > 1, ]

# Save results to file
write.csv(as.data.frame(res), "DE_results.csv")
write.csv(res, "DE_results.csv")
```

---

### 8. Additional R Packages:

```{r ch8}
# Read in more data if you want full pipeline:
library("DESeq2")

# Advanced differential expression with DESeq2
dds <- DESeq(dds)

# Visualization
plotPCA(dds, plot.number = 1)
```

---

### Notes and Suggestions:

- Adjust package usage based on your data size and computational resources.
- Use `lrm::` for large-scale analysis.
- For more complex analyses such as DESeq2, limma, or edgeR, refer to their 
documentation pages.

---

Would you like help adapting this code to a specific data type (e.g., 
single-cell, spatial transcriptomics) or integrating a specific package (e.g., 
`Seurat`, `monovar`)?
